---
title: "DGA Deep Learning"
output: html_notebook
---

```{r}
library(tidyverse)
library(dbplyr)
library(RMySQL)
library(stringr)
library(keras)
library(reshape2)
library(abind)
library(caret)
library(plotROC)
library(plotly)
```
# Database Generation
```{r create-db }
con <- DBI::dbConnect(RMySQL::MySQL(), 
  host = "localhost",
  user = "root",
  dbname="DGA",
  password = "dios"
)
domains_db <- tbl(con, "Domains")
domains_tbl<-collect(domains_db)
domains_tbl %>% filter(label!='Normal')


domains_tbl<-domains_tbl %>%  mutate(label=str_replace(label,"DGA.360","DGA")) %>% mutate(label=tolower(label))
domains_tbl<-domains_tbl %>%  mutate(label=str_replace(label,"conflicker","conficker")) %>% mutate(label=tolower(label))

labels<-domains_tbl %>% group_by(label) %>% summarise(n=n()) %>% select(label) 
labels<-as.vector(unlist(labels[1])) %>% unique()
blacklisted_labels = c('dga.banjori', 'dga.suppobox', 'dga.volatile', 'dga.matsnu', 'dga.beebone', 'dga.madmax', 'dga.cryptowall')
labels<-setdiff(labels,blacklisted_labels)
domains_tbl %>% filter (label=='normal')
```
## Dataset Class Distribution
```{r label-dist,fig.height=3, fig.width=8, paged.print=TRUE}
dga_labels_freq<-domains_tbl %>% filter(!(label %in% blacklisted_labels)) %>% group_by(label) %>% summarise(freq=n()) %>% filter(grepl('dga',label))
normal_labels_freq<-domains_tbl %>%  filter(!(label %in% blacklisted_labels)) %>% group_by(label) %>% summarise(freq=n()) %>% filter(!grepl('dga',label))


labels_freq <- domains_tbl %>% filter(!(label %in% blacklisted_labels)) %>% summarise(legit=sum(ifelse(!grepl('dga',label),1,0)),dga=sum(ifelse(grepl('dga',label),1,0)))


cbind(dga=dga_labels_freq %>% summarise(dga=sum(freq)),legit=normal_labels_freq %>% summarise(legit=sum(freq)))

```

```{r}


ggplot(melt(labels_freq ))+
  geom_col(aes(x=variable,y=value,fill=variable))+
  theme_bw()

```
## Malware families Episode Frequency Distribution
```{r}
dga_labels_freq_fig<-ggplot(dga_labels_freq  %>% arrange(desc(freq)))+
  geom_col(aes(x=label,y=freq),fill='white',color='black')+
  theme_bw()+
  ylab("Total")+
  xlab("DGA Labels")+
  theme(axis.text=element_text(size=12))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


        
ggsave(plot=dga_labels_freq_fig,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/dga_labels_freq_fig.eps",device = "eps", width =10, height = 4)

ggplot(normal_labels_freq)+
  geom_col(aes(x=label,y=freq),fill='orange')+
  theme_bw()+
  ylab("Total")+
  xlab("Normal Labels")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

dga_labels_freq_fig
```
## Top 5 DGA families
```{r top 5 DGA}
top_dga_families<-(dga_labels_freq  %>% arrange(desc(freq)) %>% top_n(5))
sum((dga_labels_freq  %>% arrange(desc(freq)) %>% top_n(5))$freq) / sum(dga_labels_freq$freq) 
top_dga_families
```
## Normal Character Frequency Distribution
```{r charlist dga}
class_dga=domains_tbl %>% filter(grepl("dga",label)) %>% filter(!(label %in% blacklisted_labels))
# calculate Char freq by DGA type
charlist_by_type=class_dga  %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))

dga_type=as.vector(unlist(charlist_by_type$charlist))

charlist_dga_plot<-ggplot(data.frame(charlist=dga_type),aes(x=charlist))+
   geom_bar(col="black",fill='black',aes(y = (..count..)/sum(..count..)))+
  scale_y_continuous(labels=percent)+ylab("Percent")+xlab("Domain Characters")+
  ggtitle("Character Frequency Distribution for DGA domains")+
  theme_bw()

ggsave(plot=charlist_dga_plot,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/dga_charlist_freq.eps",device = "eps", width =8, height = 2)
charlist_dga_plot
```

## DGA Character Frequency Distribution
```{r fig.height=4, fig.width=8}
class_normal=domains_tbl %>% filter(grepl("normal",label))
# calculate Char freq by DGA type
charlist_by_type=class_normal  %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))

normal_type=as.vector(unlist(charlist_by_type$charlist))
normal_type_letters_only<-normal_type [ !normal_type %in% c("1","2","3","4","5","6","7","8","9","0","_","-",".")]


charlist_normal_plot<-ggplot(data.frame(charlist=normal_type) %>% filter(),aes(x=charlist))+
   geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
  scale_y_continuous(labels=percent)+ylab("Percent")+xlab("Domain Characters")+
  ggtitle("Character Frequency Distribution for Normal domains")+
  theme_bw()


charlist_normal_plot_type_letters_only<-ggplot(data.frame(charlist=normal_type_letters_only) %>% filter(),aes(x=charlist))+
   geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
  scale_y_continuous(labels=percent)+ylab("Percent")+xlab("Domain Characters")+
  ggtitle("Character Frequency Distribution for Normal domains")+
  theme_bw()

ggsave(plot=charlist_normal_plot,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/normal_charlist_freq.eps",device = "eps", width =8, height = 2)
# only letters
ggsave(plot=charlist_normal_plot_type_letters_only,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/normal_charlist_freq_only_letters.eps",device = "eps", width =8, height = 2)

charlist_normal_plot
charlist_normal_plot_type_letters_only

```
# Dataset preprocessing
## Base functions 
```{r build datasets}
set.seed(12121)
train_test_sample<-function(x,percent=0.7){
  smp_size <- floor(percent * nrow(x))
  train_ind <- sample(seq_len(nrow(x)), size = smp_size)
  return (train_ind)
}
train_normal=c()
valid_normal=c()
test_normal=c()

train_dga=c()
valid_dga=c()
test_dga=c()
for (i in 1:length(labels)){
  d<-domains_tbl %>% filter(label==labels[i]) #%>% head(10000)
  d$domain1<-str_split(d$domain,"\\.",simplify = T)[,1]
  # TODO: Reemplazar por sample() 
  dindex<-train_test_sample(d,0.4)
  train<-d[dindex,]
  test<-d[-dindex,]
  tindex<-train_test_sample(test,0.5)
  valid<-test[tindex,]
  test<-test[-tindex,]
  if(str_detect(labels[i],"normal")){
    train_normal<-rbind(train_normal,cbind(domain=train$domain1,label=train$label))
    valid_normal<-rbind(valid_normal,cbind(domain=valid$domain1,label=valid$label))
    test_normal<-rbind(test_normal,cbind(domain=test$domain1,label=test$label))
  }else{
    train_dga<-rbind(train_dga,cbind(domain=train$domain1,label=train$label))
    valid_dga<-rbind(valid_dga,cbind(domain=valid$domain1,label=valid$label))
    test_dga<-rbind(test_dga,cbind(domain=test$domain1,label=test$label))
  }
}

```

```{r base-functions}
library(keras)

valid_characters <- "$abcdefghijklmnopqrstuvwxyz0123456789-_."
valid_characters_vector <- strsplit(valid_characters,split="")[[1]]
tokens <- 0:length(valid_characters_vector)
names(tokens) <- valid_characters_vector
#
# Aux functions
#

# convert dataset to matrix of tokens  
tokenize <- function(data,labels){
  #string_char_vector <- strsplit(string,split="")[[1]]
   x_data <- sapply(
     lapply(data,function(x) strsplit(tolower(x),split="")),function(x) lapply(x[[1]], function(x) tokens[[x]]))
     padded_token<-pad_sequences(x_data,maxlen=45,padding='post', truncating='post')
  return (list(encode=padded_token,domain=data, label=labels))
  #return (list(encode=padded_token))
} 
# convert vector with char tokens to one-hot encodings
to_onehot <- function(data,shape){
  train <- array(0,dim=c(shape[1],shape[2],shape[3]))
  for (i in 1:shape[1]){
    for (j in 1:shape[2])
      train[i,j,data[i,j]] <- 1
  }
  return (train)
}

# Create a dataset 
build_dataset<- function(data,maxlen){
  dataset<-tokenize(data[,1],data[,2])
  shape=c(nrow(dataset$encode),maxlen,length(valid_characters_vector))
  #dataset$encode<-to_onehot(dataset$encode,shape)
  return(dataset)
}
```
# The actual dataset creation
```{r}
maxlen=45
train_normal_dataset<-build_dataset(train_normal,maxlen)
valid_normal_dataset<-build_dataset(valid_normal,maxlen)
test_normal_dataset<-build_dataset(test_normal,maxlen)

train_dga_dataset<-build_dataset(train_dga,maxlen)
valid_dga_dataset<-build_dataset(valid_dga,maxlen)
test_dga_dataset<-build_dataset(test_dga,maxlen)



```
## Save dataset objects
```{r save dataset}

save(train_normal_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/train_normal_dataset.rda")
save(valid_normal_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/valid_normal_dataset.rda")
save(test_normal_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/test_normal_dataset.rda")

save(train_dga_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/train_dga_dataset.rda")
save(valid_dga_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/valid_dga_dataset.rda")
save(test_dga_dataset,file="/home/harpo/Dropbox/ongoing-work/git-repos/dga-algos/data/test_dga_dataset.rda")

```
## Keras Dataset Narray reshape

```{r keras data}
train_dataset_x<-abind(train_dga_dataset$encode,train_normal_dataset$encode,along=1)
train_dataset_y<-c(rep(1,nrow(train_dga_dataset$encode)),rep(0,nrow(train_normal_dataset$encode)))

valid_dataset_x<-abind(valid_dga_dataset$encode,valid_normal_dataset$encode,along=1)
valid_dataset_y<-c(rep(1,nrow(valid_dga_dataset$encode)),rep(0,nrow(valid_normal_dataset$encode)))

test_dataset_x<-abind(test_dga_dataset$encode,test_normal_dataset$encode,along=1)
test_dataset_y<-c(rep(1,nrow(test_dga_dataset$encode)),rep(0,nrow(test_normal_dataset$encode)))

```
## Keras input layer
```{r keras model}

input_shape <- dim(train_dataset_x)[2]
inputs<-layer_input(shape = input_shape) 
```
### On embedding

As far as I know, the Embedding layer is a simple matrix multiplication that transforms words into their corresponding word embeddings.

The weights of the Embedding layer are of the shape (vocabulary_size, embedding_dimension). For each training sample, its input are integers, which represent certain words. The integers are in the range of the vocabulary size. The Embedding layer transforms each integer i into the ith line of the embedding weights matrix.

In order to quickly do this as a matrix multiplication, the input integers are not stored as a list of integers but as a one-hot matrix. Therefore the input shape is (nb_words, vocabulary_size) with one non-zero value per line. If you multiply this by the embedding weights, you get the output in the shape

(nb_words, vocab_size) x (vocab_size, embedding_dim) = (nb_words, embedding_dim)
So with a simple matrix multiplication you transform all the words in a sample into the corresponding word embeddings

### On layer_conv_1d
*filters:* Integer, the dimensionality of the output space (i.e. the number output of filters in the convolution). Un filtro se corresponde a una neurona y cuyo peso determinan el kernel utilizado para la convolucion. Cada neurona tendra una convolucion diferente con un kernel distinto(el kernel es el mismo??), aunque del mismo tamaño.

*kernel_size:* An integer or list of a single integer, specifying the length of the 1D convolution window.

*strides:* An integer or list of a single integer, specifying the stride length of the convolution. Specifying any stride value != 1 is incompatible with specifying any dilation_rate value != 1.

## Keras model creation and training
```{r keras model convnet}
nb_filter <- 256
kernel_size <- 4
embedingdim <- 100

embeding<- inputs %>% layer_embedding(length(valid_characters_vector), embedingdim , input_length = input_shape)

conv1d <- embeding %>%
  layer_conv_1d(filters = nb_filter, kernel_size = kernel_size, activation = 'relu', padding='valid',strides=1) %>%
  layer_flatten() %>%
  layer_dense(1024,activation='relu') %>%
  layer_dense(1,activation = 'sigmoid')

#compile model
model <- keras_model(inputs = inputs, outputs = conv1d) 
model %>% compile(
  optimizer = 'adam',
  loss = 'binary_crossentropy',
  metrics = c('accuracy')
)

summary(model)
tensorboard("logs/run_a")
callbacks = list(
  callback_tensorboard(
    log_dir = "logs/run_a",
    histogram_freq = 1,               
    embeddings_freq = 1,
    write_graph = FALSE
  )
)


history_5epochs<-model %>% fit(train_dataset_x,train_dataset_y,epochs = 5, batch_size = 4096, validation_data = list(valid_dataset_x,valid_dataset_y))
#model %>% save_model_hdf5("cnn-nomlp.h5")

```
## Save KERAS object for futher plot
```{python eval=FALSE, include=FALSE}
from keras.utils import plot_model
from keras.models import load_model
py_model=model = load_model('cnn-5e.h5')
#plot_model(model, to_file='model.png')
```

## Keras model learning rate
```{r}
history_plot<-plot (history_5epochs)+
  #geom_line()+
  theme_bw()+
  theme(legend.position="bottom")

#ggsave(plot=history_plot,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/nn_epoch.eps",device = "eps", width =8, height = 3)
history_plot
```
## Keras model results on testset
```{r test}

predsprobs<-model %>% predict(test_dataset_x, batch_size=4096)
preds<-ifelse(predsprobs>0.9,1,0)
confusionMatrix(preds,test_dataset_y,positive = '1')
```
## Keras Model AUC
```{r roc curve}
auc_plot<-ggplot(data.frame(yes=predsprobs,class=test_dataset_y), aes(m = yes, d = factor(class, labels=c("no","yes"),levels = c(0, 1)))) + 
    geom_roc(hjust = -0.4, vjust = 1.5, n.cuts=20,colour='orange') + 
    geom_abline(intercept = 0, slope = 1,colour='red')+
    xlab("False Alarm Rate")+
    ylab("Attack Detection Rate")+
  theme_bw()
auc_plot
ggsave(plot=auc_plot,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/auc_plot.eps",device = "eps", width =8, height = 3)


 
data.frame(pred=predsprobs,class=test_dataset_y,label= c(test_dga_dataset$label,test_normal_dataset$label)) 

```
# Keras model results 
## Keras model results per DGA family
```{r analisys per DGA type, fig.height=6, fig.width=12}

nntest<-data.frame(predicted_probability=predsprobs,class=test_dataset_y,label= c(test_dga_dataset$label,test_normal_dataset$label),domain=c(test_dga_dataset$domain,test_normal_dataset$domain)) 

nntest<-nntest %>% mutate(predicted_class=ifelse(predicted_probability>0.9,1,0))
nntest<-nntest %>% mutate(label=str_replace(label,"conflicker","conficker"))
recall<-nntest %>% group_by(label) %>% summarise(recall=sum(predicted_class==class)/n(),support=n()) 
recall_dga<-recall %>% filter(!grepl('normal',label)) %>% mutate(label=str_replace(label,"conflicker","conficker"))
recall_normal<-recall %>% filter(grepl('normal',label)) %>% mutate(label=str_replace(label,"conflicker","conficker"))

#precision_far=nntest %>% group_by(label) %>% summarise(precision=1-(sum(predicted_class!=class)/n()),support=n()) 
#recall=recall %>% filter( label %in% worbased)
#confusionMatrix(nntest$predicted_class,nntest$class)

whaleboneplot<-ggplot(recall_dga %>% filter(!grepl('normal',label)) %>% mutate(label=str_replace(label,"conflicker","conficker")),aes(x=label,y=recall))+
  geom_point(aes(size=support),color='black',fill='orange',shape = 21,alpha=0.5)+
  geom_point(size=1,color='blue',fill='blue',shape = 21)+
  ylab("Attack Detection Rate")+
  ggtitle("")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_gradient2(low = "cyan", mid='blue',high = "red")+
   xlab('DGA Malware families' )+
   guides(colour=FALSE,size=FALSE)+
    scale_size_continuous(range = c(5,15))
 # scale_size_area(max_size = 15) 
ggsave(plot=whaleboneplot,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/recall_per_malware_2.png",device = "png", width =8, height = 4)

whaleboneplot  
```
## Keras model results per normal type
```{r }
whaleboneplot_normal<-ggplot(recall_normal,aes(x=label,y=1-recall))+
  geom_point(aes(size=support),color='black',fill='skyblue',shape = 21,alpha=0.5)+
  geom_point(size=1,color='blue',fill='blue',shape = 21)+
  ylab("False Positive Rate")+
  ggtitle("")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_gradient2(low = "cyan", mid='blue',high = "red")+
   xlab('Normal types' )+
  scale_x_discrete(labels=c("normal.bambenek","normal.alexa"))+
   guides(colour=FALSE,size=FALSE)+
    scale_size_continuous(range = c(5,15))
 # scale_size_area(max_size = 15) 
ggsave(plot=whaleboneplot_normal,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/recall_per_normal.png",device = "png", width =4, height = 2.5)
whaleboneplot_normal
```

## ADR vs frequency per DGA family
```{r}
recall  %>%  arrange((recall))

freq_adr_fig<-ggplot(recall,aes(x=recall,y=support))+
  geom_point(color='blue',aes(text=label),alpha=0.6)+
  xlab("Attack Detection Rate")+ylab("Frequency")+
  geom_smooth(method='lm',color='orange')+
  
  geom_point(x=0.7858304,y=30149,color='red')+
  geom_text(x=0.7858304,y=34049,label='conficker')+
  
  geom_point(x=0.6389513,y=1335,color='red')+
  geom_text(x=0.6389513,y=-1335,label='symmi')+
  
  geom_point(x=0.4976382,y=3599,color='red')+
  geom_text(x=0.4976382,y=6599, label='virut')+
  
  geom_point(x=0.9915788,y=33843,color='green')+
  geom_text(x=0.9915788,y=36843, label='cryptolocker')+
  
  geom_point(x=0.9984641,y=30600,color='green')+
  geom_text(x=0.9984641,y=27600, label='ramdo')+
  
  geom_point(x=0.9999697,y=66000,color='green')+
  geom_text(x=0.9999697,y=64000, label='post')+
  
  geom_point(x=0.9953588,y=58174,color='green')+
  geom_text(x=0.9953588,y=55174, label='tinba')+
  
  theme_bw()
freq_adr_fig
ggplotly()
recall %>% inner_join(top_dga_families,by='label')
        
ggsave(plot=freq_adr_fig,filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/dga_freq_adr_fig.png",device = "png", width =10, height = 4)


```

```{r}
nntest

FN_domains<-nntest %>% filter(predicted_class == 0 & class == 1) %>% select(domain,predicted_probability,label)
FP_domains<-nntest %>% filter(predicted_class == 1 & class == 0) %>% select(domain,predicted_probability,label)
TP_domains<-nntest %>% filter(predicted_class == 1 & class == 1) %>% select(domain,predicted_probability,label)
TN_domains<-nntest %>% filter(predicted_class == 0 & class == 0) %>% select(domain,predicted_probability,label)
```
## Character Frequency distribution of results (FN,FP,TN,TP)
```{r char distribution }
char_dist<- function(label_name){
  library(scales)
  
  charlist_FN=FN_domains %>% filter(grepl(label_name,label)) %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))
  charlist_TP=TP_domains %>% filter(grepl(label_name,label)) %>%group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))
  
  charlist_FN=as.vector(unlist(charlist_FN %>% select(charlist)))
  charlist_TP=as.vector(unlist(charlist_TP %>% select(charlist)))
  
  tpplot<-ggplot(data.frame(charlist=charlist_TP),aes(x=charlist))+
    geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
    theme_bw()
  
  
  fnplot<-ggplot(data.frame(charlist=charlist_FN),aes(x=charlist))+
    geom_bar(col="black",fill='black',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
    theme_bw()
  
  #gridExtra::grid.arrange(fnplot,tpplot,ncol=1)
  return(list(fnplot=fnplot,tpplot=tpplot))
}

conf_charlist<-char_dist("dga.conficker")
symmi_charlist<-char_dist("dga.symmi")
virut_charlist<-char_dist("dga.virut")

gridExtra::grid.arrange(
conf_charlist$fnplot+ggtitle("Conficker False Negative Character Distribution"),
symmi_charlist$fnplot+ggtitle("Symmi False Negative Character Distribution"),
virut_charlist$fnplot+ggtitle("Virut False Negative Character Distribution"),
ncol=1)

ggsave(plot=conf_charlist$fnplot+ggtitle("Conficker False Negative Character Frequency Distribution"),
       filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/conficker_charlist_freq.eps",device = "eps", width =8, height = 2)

ggsave(plot=symmi_charlist$fnplot+ggtitle("Symmi False Negative Character Frequency Distribution"),
       filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/symmir_charlist_freq.eps",device = "eps", width =8, height = 2)

ggsave(plot=virut_charlist$fnplot+ggtitle("Virut False Negative Character Frequency Distribution"),
       filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/virut_charlist_freq.eps",device = "eps", width =8, height = 2)


```
## PCA projection of results (FN, FP, TN, TP)
```{r PCA }
pca_projection<-function(label_name){
  FN_tokenized=tokenize(as.matrix(FN_domains  %>% filter(grepl(label_name,label)) %>% select(domain)),FN_domains$label)
  TP_tokenized=tokenize(as.matrix(TP_domains  %>% filter(grepl(label_name,label)) %>% select(domain)),TP_domains$label)
  FN_tokenized$label<-rep("FN", length(FN_tokenized$domain))
  TP_tokenized$label<-rep("TP", length(TP_tokenized$domain))
  
  malware_results=list()
  malware_results$encode<-abind(FN_tokenized$encode,TP_tokenized$encode,along=1)
  malware_results$domain<-c(FN_tokenized$domain,TP_tokenized$domain)
  malware_results$label<-c(FN_tokenized$label,TP_tokenized$label)
  
  nrow(malware_results$encode)
  length(malware_results$label)
  length(malware_results$domain)
  
  pca=prcomp(malware_results$encode[,1:11],center=TRUE,scale.=TRUE)
  pca_data<-data.frame(pca$x,label=malware_results$label,domain=malware_results$domain)
  pca_plot<-ggplot(pca_data %>% sample_n(1000),aes(x=PC1,y=PC4))+
    geom_point(aes(color=label,text=domain),alpha=0.5)+
    theme_bw()
  
  return (list(pca_plot=pca_plot,pca_data=pca_data))
}
r<-pca_projection("dga.conficker")
r<-pca_projection("dga.symmi")
r<-pca_projection("dga.virut")

r$pca_plot

ggsave(plot=r$pca_plot+ theme(legend.position="bottom"),
       filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/conficker_fn_pca.pdf",device = "pdf", width =8, height = 3)
#ggplotly()
#plot_ly(r$pca_data %>% sample_n(1000), type="scatter3d", x = ~PC1, y = ~PC2, z = ~PC3, color = ~label, colors = c('#BF382A', '#0C4B8E'), opacity=0.5, marker = list(size = 3),text = ~domain) %>% layout(title = "Detection Results using Convnet 1D")
```

## Character len distribution
```{r}

len_freq<-ggplot(r$pca_data  %>% mutate(len=str_length(domain)))+
  geom_histogram(aes(x=len,fill=label),bins=15)+ylab("Frequency Episode")+xlab("length")+
  theme_bw()

ggsave(plot=len_freq+ theme(legend.position="bottom"),
       filename = "/home/harpo/Dropbox/ongoing-work/publicaciones/argencon2018-DGANN/conficker_len.pdf",device = "pdf", width =8, height = 3)
len_freq
```

```{r}
charlist_FN=FN_domains %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))
charlist_FP=FP_domains %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))
charlist_TP=TP_domains %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))
charlist_TN=TN_domains %>% group_by(label) %>% do(charlist= unlist(sapply(.$domain, function(x) c(str_split(x,"")[1]))))

charlist_FN=as.vector(unlist(charlist_FN %>% select(charlist)))
charlist_FP=as.vector(unlist(charlist_FP %>% select(charlist)))
charlist_TP=as.vector(unlist(charlist_TP %>% select(charlist)))
charlist_TN=as.vector(unlist(charlist_TN %>% select(charlist)))

save(charlist_FN,file  ="/home/harpo/Dropbox/ongoing-work/git-repos/dga-wb-r/charlist_FN.rda")
save(charlist_FP,file  ="/home/harpo/Dropbox/ongoing-work/git-repos/dga-wb-r/charlist_FP.rda")
save(charlist_TP,file  ="/home/harpo/Dropbox/ongoing-work/git-repos/dga-wb-r/charlist_TP.rda")
save(charlist_TN,file  ="/home/harpo/Dropbox/ongoing-work/git-repos/dga-wb-r/charlist_TN.rda")

 
```

```{r}
library(scales)
tpplot<-ggplot(data.frame(charlist=charlist_TP),aes(x=charlist))+
    geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
   theme_bw()

tnplot<-ggplot(data.frame(charlist=charlist_TN),aes(x=charlist))+
    geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
   theme_bw()



fnplot<-ggplot(data.frame(charlist=charlist_FN),aes(x=charlist))+
    geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
   theme_bw()

fpplot<-ggplot(data.frame(charlist=charlist_FP),aes(x=charlist))+
    geom_bar(col="black",fill='white',aes(y = (..count..)/sum(..count..)))+
    scale_y_continuous(labels=percent)+ylab("Percent")+xlab("")+
   theme_bw()
gridExtra::grid.arrange(fnplot,tpplot,fpplot,tnplot,ncol=1)
```

```{r keras model lstm endgame}
input_shape <- dim(train_dataset_x)[2]
inputs<-layer_input(shape = input_shape) 

embeding<- inputs %>% layer_embedding(length(valid_characters_vector), 128 , input_length = input_shape)

lstm <- embeding %>%
  layer_lstm(units = 128) %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(1, activation = 'sigmoid')

#compile model
model_endgame <- keras_model(inputs = inputs, outputs = lstm)
model_endgame %>% compile(
  optimizer = 'rmsprop',
  loss = 'binary_crossentropy',
  metrics = c('accuracy')
)
summary(model_endgame)
model_endgame %>% fit(train_dataset_x,train_dataset_y,epochs = 10, batch_size = 4096, validation_data = list(valid_dataset_x,valid_dataset_y))

```
```{r endgame evaluation}
predsprobs_endgame<-model_endgame %>% predict(test_dataset_x, batch_size=4096)
preds_endgame<-ifelse(predsprobs_endgame>0.9,1,0)
confusionMatrix(preds_endgame,test_dataset_y,positive = '1')
```


```{r analisys per DGA type for Endgame, fig.height=4, fig.width=12}

nntest_endgame<-data.frame(predicted_probability=predsprobs_endgame,class=test_dataset_y,label= c(test_dga_dataset$label,test_normal_dataset$label),domain=c(test_dga_dataset$domain,test_normal_dataset$domain)) 

nntest_endgame<-nntest_endgame %>% mutate(predicted_class=ifelse(predicted_probability>0.9,1,0))

recall_endgame<-nntest_endgame %>% group_by(label) %>% summarise(recall=sum(predicted_class==class)/n(),support=n()) 

plot_endgame<-ggplot(recall_endgame %>% filter(!grepl('normal',label)) ,aes(x=label,y=recall))+
  geom_point(aes(size=support),color='black',fill='skyblue',shape = 21,alpha=0.5)+
  geom_point(size=1,color='blue',fill='blue',shape = 21)+
  ggtitle("NN ")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_gradient2(low = "cyan", mid='blue',high = "red")+
   xlab('Types' )+ylab("Recall")+
   guides(colour=FALSE,size=FALSE)+
    scale_size_continuous(range = c(5,15))
 # scale_size_area(max_size = 15) 
ggplotly(plot_endgame)
```

```{r fig.height=5, fig.width=10}

recall_concat <- cbind(recall, recall2=recall_endgame$recall) %>% mutate(sign=ifelse(recall-recall2>=0,'positive','negative'))

#recall_concat <- reshape2::melt(recall_contact)
ggplot(recall_concat,aes(x=label))+
  geom_col(aes(y=recall-recall2,fill=sign),alpha=0.5)+
  
  ggtitle("NN ")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_gradient2(low = "cyan", mid='blue',high = "red")+
   xlab('Types' )+ylab("Recall Diff (convnet - lstm)")+
   guides(colour=FALSE,size=FALSE)+
    scale_size_continuous(range = c(5,15))

ggplotly()

gridExtra::grid.arrange(whaleboneplot,plot_endgame,nrow=1)
```

